---
title: "exp027j2_plotting"
output: html_notebook
---


based on exp011b plotting

takes data from exp027j2_load and wrangle, exp027j2_nonlinear fitting



# libraries
```{r}
library(tidyverse)
library(broom)
library(qs)
```


# directories
```{r}
# set directory path for input data (NOTE ONLY USE CORRECTED DATA)
input_free <- "./input/"

# output
output_directory <- "./output/"


```

# import wrangled peakdata and fits
```{r}
labeled.data <- read_tsv(paste0(output_directory, "labeled data.tsv"))

fits.ssasymp <- qread(file = paste0(output_directory, "fits_ssasymp.qs")) # reading nested tables requires qs package
```

filter away some unidentified samples
```{r}
labeled.data <-
  labeled.data %>% 
    drop_na(Animal) # remove unidentified samples
  
  
fits.ssasymp <-
  fits.ssasymp %>%
    drop_na(Method) # remove unidentified samples

```

# plot plasma amino acid labeling for each animal to confirm infusions
```{r}

g <-
labeled.data %>%
  filter(isotopeLabel == "C12 PARENT") %>%
    filter(Tissue == "plasma") %>%
  ggplot(aes(x=Hours, y=(1-fraction.all), color=compound)) +
    egg::theme_article() +
    geom_point() +
    geom_line() +
    facet_wrap(~Animal)
g
ggsave(g, filename="plasma labeling from each animal.pdf", device="pdf", width=14, height=10, limitsize=T, path=output_directory)


```


filter away samples from mice which had failed infusions
```{r}

# M02 was ok but had low labeling throughout
labeled.data <- labeled.data %>% filter(Animal != "exp027j2_M02")

# M04 was good at first, but failed between 2-4 hours.
labeled.data <- labeled.data %>% filter(Animal != "exp027j2_M04")

```





# plot overall labeling

```{r}
# all compounds faceted
g <-
labeled.data %>%
  #filter(Compound %in% c("lactate", "pyruvate")) %>%
    filter(isotopeLabel != "C12 PARENT") %>%
      filter(Tissue == "plasma") %>%
      ggplot(aes(x=Hours, y=fraction.all, color=isotopeLabel, shape=Animal)) +
         egg::theme_article() +
         geom_point() +
         #ylim(0, 0.4) +
         theme(aspect.ratio = 1) +
         facet_wrap(compound~Method)
g

ggsave(g, filename="all compounds labeling.pdf", device="pdf", width=21, height=21, limitsize=T, path=output_directory)


```

simplify to 1-C12 Parent
```{r}

# all compounds faceted
g <-
labeled.data %>%
  #filter(Compound %in% c("lactate", "pyruvate")) %>%
    filter(isotopeLabel == "C12 PARENT") %>%
      filter(Tissue == "plasma") %>%
      ggplot(aes(x=Hours, y=(1-fraction.all), color=Animal)) +
         egg::theme_article() +
         geom_point() +
         geom_line(aes(group=Animal)) +
         #ylim(0, 0.4) +
         theme(aspect.ratio = 1) +
         facet_wrap(compound~Method)
g

ggsave(g, filename="all compounds labeling combined.pdf", device="pdf", width=21, height=21, limitsize=T, path=output_directory)

```


# compare labeling between tissues and plasma

```{r}
# compare Normalized Labeling (tissue/plasma) for amino acids only
# reorganized from exp011b version so that this will always plot all tissues included
g<-
labeled.data %>%
  filter(Method =="free") %>% filter(Hours==4) %>%
    filter(compound %in% c("arginine", "histidine", "lysine", "aspartate", "glutamate", "serine", "threonine", "asparagine", "glutamine", "cysteine", "selenocysteine", "glycine", "proline", "alanine", "valine", "isoleucine", "leucine", "methionine", "phenylalanine", "tyrosine", "tryptophan")) %>%
    filter(isotopeLabel == "C12 PARENT") %>%
      mutate(plasma = if_else(Tissue=="plasma", fraction.all, NA)) %>% # pull out plasma values into a new column
        group_by(compound, Hours, Tracer, Animal) %>%
          mutate(ref.plasma = max(plasma, na.rm=T)) %>% # apply plasma values to corresponding observations
        ggplot(aes(x=compound, y=(1-fraction.all)/(1-ref.plasma), color=Tissue)) +
          egg::theme_article() +
          geom_point() +
          #xlim(0, 0.15) +
          ylim(0, 2) +
          ylab("Tissue/Plasma") +
          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
          #facet_wrap(~compound) +
          ggtitle("Free Amino Acid Labeling in plasma vs tissues", subtitle="Normalized Labeling: Tissue/Plasma")
g
ggsave(g, filename="tissue vs plasma amino acids normalized labeling.pdf", device="pdf", width=14, height=8, limitsize=T, path=output_directory)


# all compounds
g<-
labeled.data %>%
  filter(Method =="free") %>% filter(Hours==4) %>%
    filter(isotopeLabel == "C12 PARENT") %>%
      mutate(plasma = if_else(Tissue=="plasma", fraction.all, NA)) %>% # pull out plasma values into a new column
        group_by(compound, Hours, Tracer, Animal) %>%
          mutate(ref.plasma = max(plasma, na.rm=T)) %>% # apply plasma values to corresponding observations
        ggplot(aes(x=compound, y=(1-fraction.all)/(1-ref.plasma), color=Tissue)) +
          egg::theme_article() +
          geom_point() +
          #xlim(0, 0.15) +
          ylim(0, 2) +
          ylab("Tissue/Plasma") +
          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
          #facet_wrap(~compound) +
          ggtitle("Free Amino Acid Labeling in plasma vs tissues", subtitle="Normalized Labeling: Tissue/Plasma")
g
ggsave(g, filename="tissue vs plasma all compounds normalized labeling.pdf", device="pdf", width=14, height=8, limitsize=T, path=output_directory)
```



scatter tissue L vs plasma L
```{r}
# similar but focus on each tissue, showing scatter of plasma vs tissue for all compounds
g<-
labeled.data %>%
  filter(Method =="free") %>%
    filter(Hours==4) %>%
    filter(compound %in% c("arginine", "histidine", "lysine", "aspartate", "glutamate", "serine", "threonine", "asparagine", "glutamine", "cysteine", "selenocysteine", "glycine", "proline", "alanine", "valine", "isoleucine", "leucine", "methionine", "phenylalanine", "tyrosine", "tryptophan")) %>%
    filter(isotopeLabel == "C12 PARENT") %>%
      mutate(plasma = if_else(Tissue=="plasma", fraction.all, NA)) %>% # pull out plasma values into a new column
        group_by(compound, Hours, Tracer, Animal) %>%
          mutate(ref.plasma = max(plasma, na.rm=T)) %>% # apply plasma values to corresponding observations
        ggplot(aes(x=(1-ref.plasma), y=(1-fraction.all), color=compound)) +
          egg::theme_article() +
          geom_point() +
          facet_wrap(~Tissue) +
          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
          geom_abline(slope=1) +
          xlim(0, 0.5) +
          ylim(0, 0.5) +
          theme(aspect.ratio=1) +
          #facet_wrap(~compound) +
          ggtitle("Free Amino Acid Labeling in plasma vs tissues", subtitle="Scatter Tissue vs Plasma")
g
ggsave(g, filename="tissue vs plasma amino acids scatter.pdf", device="pdf", width=8, height=8, limitsize=T, path=output_directory)

```



focus on aspartate and valine as examples
```{r}
labeled.data %>%
  filter(Method =="free") %>%
    filter(Hours==4) %>%
    filter(compound %in% c("aspartate", "valine")) %>%
    filter(isotopeLabel == "C12 PARENT") %>%
      mutate(plasma = if_else(Tissue=="plasma", fraction.all, NA)) %>% # pull out plasma values into a new column
        group_by(compound, Hours, Tracer, Animal) %>%
          mutate(ref.plasma = max(plasma, na.rm=T)) %>% # apply plasma values to corresponding observations
        ggplot(aes(x=(1-ref.plasma), y=(1-fraction.all), color=compound)) +
          egg::theme_article() +
          geom_point() +
          facet_wrap(~Tissue) +
          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
          geom_abline(slope=1) +
          xlim(0, 0.5) +
          ylim(0, 0.5) +
          theme(aspect.ratio=1) +
          #facet_wrap(~compound) +
          ggtitle("Free Amino Acid Labeling in plasma vs tissues", subtitle="Scatter Tissue vs Plasma")


```






