---
title: "exp027j2_fitting"
output: html_notebook
---


based on exp011b fitting.

uses data from exp027j2_load and wrangle.Rmd.

perform fits and save them to qs files (special format of rdata readable by qs package).

# libraries
```{r}
library(tidyverse)
library(broom)
library(qs)
```


# directories
```{r}
# set directory path for input data (NOTE ONLY USE CORRECTED DATA)
input_free <- "./input/"

# output
output_directory <- "./output/"

```

# import wrangled data
```{r}

labeled.data <- read_tsv(paste0(output_directory, "labeled data.tsv"))

```

# fit data to curves to estimate kinetics

fit to nls starting from first timepoint (not zero) because primed injection would mess up kinetics

in theory, general fit for all metabolites should indicate water kinetics?

fitting code based on proteomics fitting code

## define prep.nls
```{r}

# prep.nls definition

# ASSUMES WORKING WITH SINGLE ISOTOPE


prep.nls <- function(df){
  df %>%
    mutate(mod.hours = Hours-4) %>%
      select(compound, Method, Hours, Tissue, Tracer, isotopeLabel, fraction.all) %>% # select relevant data
        group_by(compound, Method, Tissue, Tracer, isotopeLabel) %>% # group to key data for model
          nest()
}

# test prep.nls
labeled.data %>% filter(isotopeLabel == "C12 PARENT") %>%
    prep.nls() %>% head(5) %>% view() %>% .$data # optionally comment this in/out to preview data tables
```

## define fit functions
```{r}
# self starting assymptotic fit
ssasympit <- function(df){
  df %>%
    mutate(fit = map(data,
                     possibly(~nls(fraction.all~SSasymp(Hours, Asym, R0, lrc),
                                   data= .)
                              , otherwise="failed to fit", quiet=FALSE)),
          tidied = map(fit, possibly(tidy, otherwise="failed to tidy")),
          augmented = map(fit, possibly(augment, otherwise="failed to augment")),
    )
}
```

## run fits
```{r}
fits.ssasymp <-
  labeled.data %>%
    filter(Hours != 0) %>%
      prep.nls() %>%
        ssasympit()
```


# save

summarize first
```{r}
fits.ssasymp %>%
  filter(fit != "failed to fit") %>%
    select(compound, Method, Tracer, Tissue, isotopeLabel) %>%
      group_by(Method, Tracer, Tissue, isotopeLabel) %>%
        summarise(n.fit.compounds = n())
```


save fits df to external file (r data file bc it has nested data)
```{r}
# save(fits.ssasymp, file="fits_ssasymp.rds")
qsave(fits.ssasymp, paste0(output_directory, "fits_ssasymp.qs"))  # saving nested tables requires qs package

```





























